---
title: "Assignment 2"
author: "Jonathan Lee"
date: "September 23rd, 2020"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r, results = 'hide', message = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(magrittr)
library(tidycensus)
library(cartogram)
library(wesanderson)

```

## Baseline Map
### This is the baseline map of the US and its states using the R naturalearth package. I chose to not show the states Alaska and Hawaii on this map and the following examples.


```{r}
US_states <- ne_states(country = "United States of America",
                       returnclass = "sf") %>%
  filter(name != "Alaska",
         name != "Hawaii")


US_map <- ggplot(US_states) +
  geom_sf() +
  theme_bw()

US_map

```

## Map Projection #1
### This map is a projection using NAD83 Albers Equal Area of California.

```{r}

CA_state_plane <- "+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

US_map <- ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = CA_state_plane)+
  theme_bw()

US_map


```

## Map Projection #2
### This map is a projection using the NAD83 of Texas.

```{r}

TX_state_plane <-"+proj=lcc +lat_1=27.41666666666667 +lat_2=34.91666666666666 +lat_0=31.16666666666667 +lon_0=-100 +x_0=1000000 +y_0=1000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

US_map <- ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = TX_state_plane)+
  theme_bw()

US_map

```

## Map Projection #3
### This map is a projection using the NAD83 of Massachusetts

```{r}
MA_state_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

US_map <- ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = MA_state_plane)+
  theme_bw()

US_map

```

## Map Projection #4
### This is a projection of the US using Albers Equal Area conical projection of Brazil. You can see the conical warping of the US.


```{r}
brazil <- "+proj=aea +lat_1=-27 +lat_2=-34 +lat_0=-24 +lon_0=-54 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"

US_map <- ggplot(US_states) +
  geom_sf() +
  coord_sf(crs = brazil)+
  theme_bw()

US_map


```





## Map Projection #5
### For the purposes of the next few examples, I will use the projection of the US as a whole using the Albers Equal Area projection.

```{r}

USA_AEA <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=clrk66 +units=m +no_defs"

US_transformed <- US_states %>%
  st_transform(crs = USA_AEA)

ggplot(US_transformed) +
  geom_sf()+
  theme_bw()

```

## Chloropleth Map
### This map shows the total population of each state from the 2018 American Community Survey results. 


```{r, results = 'hide', message = FALSE}

census_data <- get_acs(geography = "state", variables = "B01003_001", year = 2018)


US_data <- left_join(US_transformed, census_data, by = c("name" = "NAME"))

ggplot(US_data, aes(fill = estimate/1000000)) +
  geom_sf()+
  scale_fill_gradientn(name = "Total population (M)",
                       breaks = seq(0,40, by = 5),
                       labels = seq(0,40, by = 5),
                       colors = wes_palette(name = "Zissou1", type = "continuous"))+
  theme_map()

```

## Cartogram Map
### This map shows the total population of each state and the distortion to its size depending on the population. 

```{r, results = 'hide', message = FALSE}

US_cartogram <- cartogram_cont(US_data, "estimate", itermax = 5)

ggplot(US_cartogram, aes(fill = estimate/1000000))+
  geom_sf(color = NA)+
  scale_fill_gradientn(name = "Total population (M)",
                       breaks = seq(0,40, by = 5),
                       labels = seq(0,40, by = 5),
                       colors = wes_palette(name = "Zissou1", type = "continuous"))+
  theme_map()


```

## Dorling Map
### This map shows the size of the population of each state as its own bubble in the location of the state.


```{r}
US_dorling <- cartogram_dorling(US_data, "estimate")

ggplot(US_dorling, aes(fill = estimate/1000000))+
  geom_sf(color = NA)+
  scale_fill_gradientn(name = "Total population (M)",
                       breaks = seq(0,40, by = 5),
                       labels = seq(0,40, by = 5),
                       colors = wes_palette(name = "Zissou1", type = "continuous"))+
  theme_map()

```

## Centroid Map
### This map shows the size of the population of each state in the center of the state's geographical location.


```{r, results = FALSE, warning = FALSE}
US_centroids <- st_centroid(US_data)

ggplot(US_data) +
  geom_sf(fill = NA, color = "gray")+
  geom_sf(data = US_centroids, aes(size = estimate/1000000),
          alpha = 0.5, color = "red") +
  scale_size_continuous(name = "Total population (M)",
                        range = c(1,20),
                        breaks = seq(0,40, by = 5),
                        labels = seq(0,40, by = 5)) +
  theme_void()
```

