---
title: "Assignment5"
author: "Jonathan Lee"
date: "October 21, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(raster)
library(tidyverse)
library(RStoolbox)
library(ggthemes)
library(ggspatial)
library(osmdata)
library(sf)
```

## Introduction
For this exercise I wanted to see how the concept of "my neighborhood" changes depending on the mode of transportation that I use. Back when I was living in LA, the spaces I would frequently go to were divided into whether I needed to drive there or if I could walk there. Therefore I drew two maps each showing the places I frequented in different modes of transportation. 


The data that I will be using for Los Angeles. Sourced from the LA Open Data portal.

```{r, message=FALSE, results='hide'}
roads <- st_read(file.path("tl_2019_06037_roads","tl_2019_06037_roads.shp"))
parcels <- st_read(file.path("la_parcels","LA_City_Parcels selection.shp"))

```

## Image 1

For this image, I drew the neighborhood I used to live in Los Angeles and the places I would frequently go to that were about 10-15 min walk from my home. This image was georeferenced in ArcGIS Pro and imported into R.

```{r}

map_image_1 <- stack("georeferencing_map_1.tif")

ggRGB(map_image_1, r = 1, g = 2, b = 3)
```

I overlayed parcel data onto my hand-drawn map.


```{r}
ggplot(parcels) +
  ggRGB(map_image_1, 
        r = 1, g = 2, b = 3, ggLayer = TRUE) +
  geom_sf(alpha = 0.1, color = "gray60") +
  coord_sf(xlim = c(map_image_1@extent@xmin, map_image_1@extent@xmax),
           ylim = c(map_image_1@extent@ymin, map_image_1@extent@ymax)) +
  theme_map() +
  annotation_scale() +
  annotation_north_arrow(location = "tr")
```

## Image 2

For this image, I drew the places I would go in about a 10-15 min car drive in Los Angeles. This image was georeferenced in ArcGIS Pro and imported into R.


```{r}
map_image_2 <- stack(file.path("C:","Users","yedah","Desktop","Fall 2020","VIS","georeferencing_map_2.tif"))

ggRGB(map_image_2, r = 1, g = 2, b = 3)
```

Then, I overlayed the roads data onto my hand-drawn map.


```{r}
ggplot(roads) +
  ggRGB(map_image_2, 
        r = 1, g = 2, b = 3, ggLayer = TRUE) +
  geom_sf(alpha = 0.5, color = "gray50") +
  coord_sf(xlim = c(map_image_2@extent@xmin, map_image_2@extent@xmax),
           ylim = c(map_image_2@extent@ymin, map_image_2@extent@ymax)) +
  theme_map() +
  annotation_scale() +
  annotation_north_arrow(location = "tr")
```
