---
title: "Assignment3"
author: "Jonathan Lee"
date: "September 29th, 2020"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, message=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(readxl)
library(gridExtra)
```


### Data reading and cleaning

```{r, message=FALSE, cache=TRUE}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

police_dist <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/9a3a8c427add450eaf45a470245680fc_5.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

community_centers <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/acedd06b39854088b0f2fefccffebf84_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D", quiet = TRUE)

crime_raw <- st_read("https://data.boston.gov/dataset/6220d948-eae2-4e4b-8723-2dc8e67722a3/resource/12cb3883-56f5-47de-afa5-3b1cf61b257b/download/tmpoyezvlyu.csv", quiet = TRUE)

nhoods_demo <- read_excel("data/nhoods_demographics.xlsx", sheet = "Age", skip = 2) %>%
  drop_na() %>%
  rename("Name" = "...1", "tot_pop" = "Total Population") %>%
  select("Name", "tot_pop", "Median Age")

crime_2020 <- crime_raw %>%
  filter(YEAR == 2020) %>%
  na_if("") %>%
  drop_na(Lat | Long)

crime_2020_sf <- st_as_sf(crime_2020, coords = c("Long", "Lat"), crs = 4326) %>%
  select(OFFENSE_CODE, OFFENSE_DESCRIPTION, geometry)

nhoods <- nhoods %>%
  left_join(nhoods_demo, by = c("Name" = "Name") )


```


### Projecting the maps onto the MA plane.

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

nhoods <- nhoods %>%
  st_transform(MA_state_plane)

community_centers <- community_centers %>%
  st_transform(MA_state_plane)

crime_2020_sf <- crime_2020_sf %>%
  st_transform(MA_state_plane)

police_dist <- police_dist %>%
  st_transform(MA_state_plane)

crime_2020_assault <- crime_2020_sf %>%
  filter(OFFENSE_CODE == "801") # Simple assault

crime_2020_auto <- crime_2020_sf %>%
  filter(OFFENSE_CODE == "723" | OFFENSE_CODE == "724") # Attempted auto theft or auto theft

```


## Map 1

This map shows the location of all auto thefts and assaults that occurred in 2020 so far.

```{r}
ggplot() +
  geom_sf(data = nhoods, fill = "gray80", color = "white") +
  geom_sf(data = crime_2020_auto, aes( color = "black"), alpha = 0.5, size =1) +
  geom_sf(data = crime_2020_assault, alpha = 0.1, aes( color = "red"))+
  theme_map()+
  annotation_scale(bar_cols = c("gray80", 'white'), pad_x = unit(7, "cm"))+
  scale_color_manual(values = c("black","red"), labels = c("Auto theft","Assault"),
                     name = "Crime")
```

## Map 2

These maps compare the difference of auto theft and assault by neighborhoods.

```{r fig.height=10, fig.width=15, message=FALSE, results='hide'}

nhoods <- nhoods %>%
  mutate(num_auto = lengths(st_covers(nhoods, crime_2020_auto))) %>%
  mutate(num_assault = lengths(st_covers(nhoods, crime_2020_assault)))


plot_auto <- ggplot(nhoods)+
  geom_sf(color = "white", aes(fill = num_auto))+
  scale_fill_gradient(name = "Number of auto theft",
                      low = "gray80", high = "#FB8C65")+
  theme_map()+
  theme(legend.position = c(0.6,0))

plot_assault <- ggplot(nhoods)+
  geom_sf(color = "white", aes(fill = num_assault))+
  scale_fill_gradient(name = "Number of Assault",
                      low = "gray80", high = "#B4281E")+
  theme_map()+
  theme(legend.position = c(0.6,0))

grid.arrange(plot_auto, plot_assault, nrow = 1)


```

## Map 3

These maps compare the difference in auto theft when counted by neighborhoods or by police districts.

```{r, fig.width=15, fig.height=10}

police_dist <- police_dist %>%
  mutate(num_auto = lengths(st_covers(police_dist, crime_2020_auto))) %>%
  mutate(num_assault = lengths(st_covers(police_dist, crime_2020_assault)))

plot_nhoods <- ggplot(nhoods)+
  geom_sf(color = "white", aes(fill = num_auto))+
  scale_fill_gradient(name = "Number of auto theft\nby neighborhood",
                      low = "gray80", high = "gray10",
                      limits = c(0,300))+
  theme_map()+
  theme(legend.position = c(0.6,0))

plot_police <- ggplot(police_dist)+
  geom_sf(color = "white", aes(fill = num_auto))+
  scale_fill_gradient(name = "Number of auto theft\nby police district",
                      low = "gray80", high = "gray10",
                      limits = c(0,300))+
  theme_map()+
  theme(legend.position = c(0.6,0))

grid.arrange(plot_nhoods, plot_police, nrow = 1)


```


## Map 5

These maps compare the difference between auto theft and assaults by police districts.

```{r, fig.width=15, fig.height=10}


plot_auto <- ggplot(police_dist)+
  geom_sf(color = "white", aes(fill = num_auto))+
  scale_fill_gradient(name = "Number of auto theft",
                      low = "gray80", high = "#2D4441")+
  theme_map()+
  theme(legend.position = c(0.6,0))

plot_assault <- ggplot(police_dist)+
  geom_sf(color = "white", aes(fill = num_assault))+
  scale_fill_gradient(name = "Number of Assault",
                      low = "gray80", high = "#8CB3AF")+
  theme_map()+
  theme(legend.position = c(0.6,0))

grid.arrange(plot_auto, plot_assault, nrow = 1)


```


## Map 6

These maps compares the area density of auto thefts when calculated by neighborhoods or police districts.

```{r, fig.width=15, fig.height=10}
nhoods <- nhoods %>%
  mutate (area = set_units(st_area(nhoods), km^2)) %>%
  mutate (auto_area_dens = as.numeric(num_auto/area))

police_dist <- police_dist %>%
  mutate (area = set_units(st_area(police_dist), km^2)) %>%
  mutate (auto_area_dens = as.numeric(num_auto/area))

plot_auto <- ggplot(nhoods)+
  geom_sf(color = "white", aes(fill = auto_area_dens))+
  scale_fill_gradient(name = "Area density of auto theft\nby neighborhoods (/km^2)",
                      low = "gray80", high = "gray10",
                      limits = c(0,40))+
  theme_map()+
  theme(legend.position = c(0.6,0))

plot_assault <- ggplot(police_dist)+
  geom_sf(color = "white", aes(fill = auto_area_dens))+
  scale_fill_gradient(name = "Area density of auto theft\nby police districts (/km^2)",
                      low = "gray80", high = "gray10",
                      breaks = seq(0,40, by=10),
                      limits = c(0,40))+
  theme_map()+
  theme(legend.position = c(0.6,0))

grid.arrange(plot_auto, plot_assault, nrow = 1)


```

## Map 7

This map shows the population density of auto thefts by neighborhoods

```{r}

nhoods <- nhoods %>%
  mutate(auto_pop_dens = as.numeric(num_auto/(tot_pop/1000))) %>%
  mutate(assault_pop_dens = as.numeric(num_assault/(tot_pop/1000)))

ggplot(nhoods)+
  geom_sf( aes(fill = auto_pop_dens, color = ""))+
  scale_fill_gradient(name = "Population density of auto theft\nby neighborhoods",
                      low = "gray90", high = "#4C4C65", na.value = "white",
                      breaks = breaks <- seq(0,3, by=1),
                      labels = paste(breaks,"auto theft per 1000 people"))+
  theme_map()+
  theme(legend.position = c(0.6,0))+
  scale_color_manual(values = NA, name = "No data")+
  guides(color=guide_legend("No data", override.aes=c(fill="white", color = "black")))



```

## Map 8

This map shows the population density of assaults by neighborhoods

```{r}
ggplot(nhoods)+
  geom_sf(aes(fill = assault_pop_dens, color = ""))+
  scale_fill_gradient(name = "Population density of assault\nby neighborhoods",
                      low = "gray90", high = "#5AB6DB", na.value = "white",
                      breaks = breaks <- seq(0,15, by=5),
                      labels = paste(breaks,"assaults per 1000 people"))+
  theme_map()+
  theme(legend.position = c(0.6,0))+
  scale_color_manual(values = NA, name = "No data")+
  guides(color=guide_legend("No data", override.aes=c(fill="white", color = "black")))

```


## Map 9

This map shows all the community centers around Boston and how far each of them are to the nearest assault.


```{r, message = FALSE, results = 'hide'}
community_centers <- community_centers %>%
  mutate(assault_dist = st_nn(community_centers, crime_2020_assault, returnDist = TRUE)$dist) %>%
  mutate(assault_dist = as.numeric(assault_dist))

avg_assault_dist <- mean(community_centers$assault_dist)


left_side <- st_bbox(nhoods)$xmin
top_side <- st_bbox(nhoods)$ymax

ggplot(nhoods) +
  geom_sf(color = "white") +
  geom_sf(data = community_centers, aes(color = assault_dist), size = 5)+
  theme_map()+
  annotate(geom = "text", x = left_side + 300, y = top_side - 550,
           label = paste("On average, the nearest distance of an assault from\n",
                         "a community center is", prettyNum(avg_assault_dist, digits = 3)),
           hjust = 0, vjust = 0, size = 3)+
  scale_color_gradient(name = "Distance of an assult\nfrom each community center",
                       low = "white", high = "#2D4441")+
  theme(legend.position = c(0.7,0))
  
```

## Map 10
This map shows the community centers and the assaults that occurred within 50m of community centers.

```{r}

community_centers_buffer <- st_buffer(community_centers, dist = 50) %>%
  st_union()

assault_community_center <- crime_2020_assault[community_centers_buffer,]

crime_assault_cc <- crime_2020_assault %>%
  st_join(assault_community_center)

crime_assault_cc <- crime_assault_cc %>%
  mutate(by_community_center = !is.na(crime_assault_cc$OFFENSE_CODE.y))

n_assault_by_cc <- sum(crime_assault_cc$by_community_center)

n_assault <- length(crime_assault_cc$by_community_center)

pct_assault <- n_assault_by_cc/n_assault * 100


ggplot(nhoods)+
  geom_sf(color = "white")+
  geom_sf(data = community_centers_buffer, aes(color = "black"), size = 5)+
  geom_sf(data = assault_community_center,
          aes(color = "red"), size = 5, alpha = 0.5)+
  scale_color_manual(values = c("black", "red"),
                     name = "Community centers and assaults within 50m",
                     labels = c("Community centers", "Assaults"))+
  annotate(geom = "text", x = left_side + 300, y = top_side - 550,
           label = paste("Of the ", prettyNum(n_assault, big.mark = ","),
                         " assaults in 2020\n",
                         prettyNum(n_assault_by_cc),
                         "(",
                         prettyNum(pct_assault, digits = 1),
                         "%) are within 50 meters of a community center."),
           hjust = 0, vjust = 0, size = 3)+
  theme_map()+
  theme(legend.position = c(0.4,0))

```

